---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# QOI

QOI is fast. It losslessy compresses images to a similar size of PNG, while offering 20x-50x faster encoding and 3x-4x faster decoding.

QOI is simple. The reference en-/decoder fits in about 300 lines of C. The file format specification is a single page PDF.

https://qoiformat.org/

This **R**-package is a wrapper to handle QOI files with **R**.
The package is completely written in base R with no dependencies and pure C-code.

<!-- badges: start -->
<!-- badges: end -->


```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "##",
                      fig.retina = 2,
                      fig.align = "center",
                      fig.path = "README_figs/README-")
Sys.setlocale("LC_TIME", "C")
```

## Installation

Until now the package is not on CRAN, so just install it via GitHub:

You can install the development version from GitHub with the following command:

```{r, eval = FALSE}
if (!require("devtools"))
  install.packages("devtools")
devtools::install_github("JohannesFriedrich/qoi4R")
```


## Usage

There are just two main functions: `readQOI()` and `writeQOI()`.
* `readQOI()`: Takes an qoi-format image and decodes it into its RGB or RGBA values. The result is a matrix with dimensions height x width x channels.
* `writeQOI()`: Takes an RGB(A) matrix and encodes it into an qoi-image.

### `readQOI()`

Let´s read in an QOI-image delivered with this package:

```{r}
library(qoi)
path <- system.file("extdata", "Rlogo.qoi", package="qoi")

Rlogo_pixels <- readQOI(path)
```

The dimension of `Rlogo_pixels` is `r dim(Rlogo_pixels)`. So the height of the image is `r dim(Rlogo_pixels)[1]` px, the width `r dim(Rlogo_pixels)[2]` px and it has `r dim(Rlogo_pixels)[3]` channels, so it´s RGBA. With `3` channels it would be RGB encoded.

With `readQOI()` is it possible to show the image with **R**.

```{r}
plot.new()
rasterImage(Rlogo_pixels/255, xleft = 0, xright = 1,
            ytop = 0, ybottom = 1, interpolate = FALSE)
```

The pixels returned from `readQOI()` are integer values between 0 - 255. To show them with `rasteImage()` it is necessary to divide them by 255 to get values between 0-1.

Let´s compare the result to the wonderful [PNG-package](https://cran.r-project.org/web/packages/png).
The results from `readPNG()` are numerical values between 0 and 1. To compare the results multiply with 255 and change the mode to integer.

```{r}
qoi_logo_rgb_png <- png::readPNG(path <- system.file("extdata", "qoi_logo.png", package="qoi"))*255

qoi_logo_rgb_qoi <- qoi::readQOI(path <- system.file("extdata", "qoi_logo.qoi", package="qoi"))

mode(qoi_logo_rgb_png) <- "integer"

identical(qoi_logo_rgb_png, qoi_logo_rgb_qoi)
```

### `writeQOI()`

With this function it is possible to save an RGB(A) matrix to an QOI-image.
The input arguments are:
* image: a matrix with dimensions height x width x channels
* target: Either name of the file to write, a binary connection or a raw vector indicating that the output should be a raw vector (so the hex-interpretation of the image).

If no second argument is given, the returned value is the hex-interpretation of the image in the QOI-file format.

```{r}
qoi_logo <- writeQOI(qoi_logo_rgb_qoi)
rawToChar(head(qoi_logo))
```


If an second argument is given as character the image is saved to this name:

```{r, eval = FALSE}
writeQOI(qoi_logo_rgb_qoi, "qoi_logo_rgb.qoi")
```

If the second argument is of type `connection` the hex interpretation of the QOI-image will be send to this connection.

```{r, eval = FALSE}
file <- file("file.qoi", "wb")
writeQOI(qoi_logo_rgb_qoi, file)
close(file)
```

## Acknowlegment 

This package would not exist without the following persons/homepages/tutorial/...:

* [Phoboslab - original specification](https://github.com/phoboslab/qoi)
* [PNG R-package](https://github.com/s-u/png)
* [Example C-code R-package](https://github.com/coolbutuseless/simplecall)
* [R extensions](https://cran.r-project.org/doc/manuals/r-release/R-exts.html)
* [Hadley´s R-Internals](https://github.com/hadley/r-internals)
